---
title: "Effect size calculation for meta-analysis"
author: "Justin Tiano"
date: "01/08/2022"
output:
  html_document:
    toc: true
    toc_float: true
---

# Introduction 

This script organizes the meta-analysis data to calculate 3 types of treatment effect sizes:
  1. Log response ratios (lnRR),
  2. Standardized mean difference (Cohen's d), and 
  3. Standardized mean difference (Hedges g). 
The resulting dataframe is saved as "Exp-Obs.rda".

```{r loading required packages, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Read data

```{r}

df <- get(load("../dataset/dataset.rdata"))

# corrections
df$Study.type[df$article_id==12]="Before After"  
df$Study.type[df$article_id==23]="Before After"
df$Study.type[df$article_id==164]="Before After"
df$Study.type[df$article_id==132]="Before After"
df$Study.type[149:150] = "Control Impact"        # Change study type for 2 parameters in article 66

# NOT DONE YET df = df[-c(197,226), ]

```

# Organize columns to calculate lnRR


```{r}

### Before After studies ###
df$response1[df$Study.type=="Before After"]=df$Mean_Before_impact[df$Study.type=="Before After"]
df$SD1[df$Study.type=="Before After"]=df$SD_Before_impact[df$Study.type=="Before After"]
df$n1[df$Study.type=="Before After"]=df$n_Before_impact[df$Study.type=="Before After"]

df$response2[df$Study.type=="Before After"]=df$Mean_After_impact[df$Study.type=="Before After"]
df$SD2[df$Study.type=="Before After"]=df$SD_After_impact[df$Study.type=="Before After"]
df$n2[df$Study.type=="Before After"]=df$n_After_impact[df$Study.type=="Before After"]

### Control Impact studies ###
df$response1[df$Study.type=="Control Impact"]=df$Mean_After_control[df$Study.type=="Control Impact"]
df$SD1[df$Study.type=="Control Impact"]=df$SD_After_control[df$Study.type=="Control Impact"]
df$n1[df$Study.type=="Control Impact"]=df$n_After_control[df$Study.type=="Control Impact"]

df$response2[df$Study.type=="Control Impact"]=df$Mean_After_impact[df$Study.type=="Control Impact"]
df$SD2[df$Study.type=="Control Impact"]=df$SD_After_impact[df$Study.type=="Control Impact"]
df$n2[df$Study.type=="Control Impact"]=df$n_After_impact[df$Study.type=="Control Impact"]

#### Before After Control Impact studies ###
df$AftCtrl_BefCtrl=df$Mean_After_control/df$Mean_Before_control
df$AftImpt_BefImpt=df$Mean_After_impact/df$Mean_Before_impact
df$SD1[df$Study.type=="Before After Control Impact"]=df$SD_After_control[df$Study.type=="Before After Control Impact"]
df$SD2[df$Study.type=="Before After Control Impact"]=df$SD_After_impact[df$Study.type=="Before After Control Impact"]
df$response1[df$Study.type=="Before After Control Impact"]=df$AftCtrl_BefCtrl[df$Study.type=="Before After Control Impact"]
df$response2[df$Study.type=="Before After Control Impact"]=df$AftImpt_BefImpt[df$Study.type=="Before After Control Impact"]
df$n1[df$Study.type=="Before After Control Impact"]=(df$n_After_control[df$Study.type=="Before After Control Impact"] 
                                                           + df$n_Before_control[df$Study.type=="Before After Control Impact"])/2
df$n2[df$Study.type=="Before After Control Impact"]=(df$n_After_impact[df$Study.type=="Before After Control Impact"] 
                                                           + df$n_Before_impact[df$Study.type=="Before After Control Impact"])/2

# Make all SD's positive
df$SD1 = abs(df$SD1)
df$SD2 = abs(df$SD2)

```

# Calculate lnRR and VlnRR

Log response ratio is calculated as (1) :
$ln(\overline{x}_{2}/\overline{x}_{1})$                                      (1), 
with $\overline{x}_{2}$ and $\overline{x}_{1})$ the means of the studied variable 
in the impact vs. control, or after vs. before samples respectively.

VarLnRR is calculated by first calculating the pooled standard deviation $S$ (2),
$S_{pooled} = \sqrt{\frac{(n_{1}-1)S^2_{1} + (n_{2}-1)S^2_{2}}{n_{1}+n_{2}-2}}$ (2)
and using $S$ to calculate the variance $V_{D}$ (3). See Borenstein et al., 2009
$V_{D} = \sqrt{\frac{S^2_{1}}{n_{1}} + \frac{S^2_{2}}{n_{2}}}$               (3)

```{r, warning=FALSE}

# Calculate lnRR
df$lnRR=log(df$response2/df$response1)

df$lnRR[df$lnRR==Inf]=NA   # anything #NAME?, Inf, and NA it replaces with NA; get rid of responses where either C or I is 0
df$lnRR[df$lnRR==-Inf]=NA  # get rid of responses where either C or I is 0

df$lnRR[df$lnRR==Inf]=NA   # get rid of responses where either C or I is 0
df$lnRR[df$lnRR==-Inf]=NA  # get rid of responses where either C or I is 0

# Pooled standard deviation
df$SD1 = as.numeric(df$SD1) # need to convert this column to numeric
df$Spooled=sqrt(((df$n1-1)*(df$SD1^2)+ (df$n2-1)*(df$SD2^2))   / # Pooled within groups SD (S within or S pooled)
              (df$n1+df$n2-2))

# VlnRR
df$VarLnRR=df$Spooled^2*( 1/(df$n1*df$response1^2) +1/(df$n2*df$response2^2))

## Before After Control Impact VarlnRR
df$VarLnRR[df$Study.type=="Before After Control Impact"]=
df$SD_After_control[df$Study.type=="Before After Control Impact"]^2/(df$n_After_control[df$Study.type=="Before After Control Impact"]*df$Mean_After_control[df$Study.type=="Before After Control Impact"]^2)+
df$SD_Before_control[df$Study.type=="Before After Control Impact"]^2/(df$n_Before_control[df$Study.type=="Before After Control Impact"]*df$Mean_Before_control[df$Study.type=="Before After Control Impact"]^2)+
df$SD_After_impact[df$Study.type=="Before After Control Impact"]^2/(df$n_After_impact[df$Study.type=="Before After Control Impact"]*df$Mean_After_impact[df$Study.type=="Before After Control Impact"]^2)+
df$SD_Before_impact[df$Study.type=="Before After Control Impact"]^2/(df$n_Before_impact[df$Study.type=="Before After Control Impact"]*df$Mean_Before_impact[df$Study.type=="Before After Control Impact"]^2)
df$SE=sqrt(df$VarLnRR)

# Weight (W) for lnRR effect size
df$W = 1/df$VarLnRR
df$W[df$W==Inf]=NA  
df$WY = df$W*df$lnRR        # Effect size multiplied by weight
df$WY2 = df$W*df$lnRR^2     # W * Y^2
df$WYsquared = df$WY^2

# Cohen's d calculations (Standardized mean difference)
# variance calculated with escalc
df <- escalc(measure="SMD", m1i=response2, m2i=response1, sd1i=SD2, sd2i=SD1, n1i=n2, n2i=n1, data=df, var.names=c("d","Vd"), digits=4) 
df$SEd = sqrt(df$Vd)
df$Wd = 1/df$Vd             # Weight for Cohen's d fixed effects
df$Wd[df$Wd==Inf]=NA  
df$WYd = df$Wd*df$d         # Cohen's d multiplied by weight
df$WYd2 = df$Wd * df$d ^ 2  # W * Y^2
df$WYdsquared = df$WYd^2

#Convert to hedges g
df$g = df$d * (1 - (3 / 4 * (df$n1 + df$n2 - 2) - 1))
df$Vg = 1 - (3 / 4 * (df$n1 + df$n2 - 2) - 1) * df$Vd
df$SEg = sqrt(df$Vg)
df$Wg = 1/df$Vg             # Weight for Hedges g fixed effects
df$Wg[df$Wg==Inf]=NA  
df$WYg = df$Wg*df$g         # Hedges g multiplied by weight
df$WYg2 = df$Wg * df$g ^ 2  # W * Y^2
df$WYgsquared = df$WYg^2

# save stuff 
save(file = "../r_objects/Exp-Obs_181022.rda", df)

```

